---
title: Arachnophobia Power analysis
format:
  html:
    toc: true
    code-tools: true
    code-line-numbers: true
    code-copy: true
    code-block-border-left: true
  typst:
    toc: true
bibliography: references.bib
lightbox: true
---

:::{.callout-note}
This example is drawn from The Babraham Institute's Statistical Analysis using R course [@carrStatisticalAnalysisUsing2025].
:::

# Packages

```{r packages}
suppressPackageStartupMessages({
	library(dplyr)
	library(readr)
	library(ggplot2)
	library(tidyr)
	library(gt)
})
```


Based on our preliminary data how many participants do we need to test if exposure to a real tarantula or pictures of that tarantula results in a more anxiety in arachnophobes?

# Data

```{r data}
spider_anxiety_data <- read_csv("data/spider-data.csv", col_types = "ffi")
spider_anxiety_data %>% gt()
```

In our data each participant has an anxiety score for each exposure,
so our experiment has a paired design.

# Exploratory Visualisation

```{r exploratoryVisualisation}
spider_anxiety_data %>%
	ggplot(aes(Group, Scores)) +
		geom_line(aes(group = Participant)) +
		geom_point(aes(colour = Participant), size = 3) +
		labs(
			title = "Anxiety Score by Exposure Type",
			x = "Exposure"
		) +
		theme_minimal()
```

# Summary Statistics

```{r summaryStats}
spider_anxiety_summary_stats <- spider_anxiety_data %>%
	group_by(Group) %>%
	summarise(mean=mean(Scores), sd=sd(Scores))

spider_anxiety_summary_stats %>% gt()
```

# Power Analysis

To perform a power analysis on pair data we need the differences between the groups.

```{r powerAnalysisDifferences}
spider_anxiety_summary_differences <- spider_anxiety_data %>%
  pivot_wider(names_from = Group, values_from = Scores) %>%
  mutate(difference = Picture - Real) %>%
  summarise(mean = mean(difference), sd = sd(difference))

spider_anxiety_summary_differences %>% gt() 
```

We will be using a simple t-test to analyses this data.

```{r powerAnalysis}
significance_threshold <- 0.05
power_level <- 0.8
t_test_type <- "paired"
t_test_directionality <- "two.sided"

spider_anxiety_t_test_power <- power.t.test(
	delta = spider_anxiety_summary_differences$mean,
	sd = spider_anxiety_summary_differences$sd,
	sig.level = significance_threshold,
	power = power_level,
	type = t_test_type,
	alternative = t_test_directionality 
)

spider_anxiety_t_test_power 
```

We are looking for a power of `r power_level * 100`% at a p<`r significance_threshold` significance threshold.
Based on this power calculation we therefore need `r ceiling(spider_anxiety_t_test_power$n)` participants to achieve the desired power.

# Session Info

```{r sessionInfo}
sessionInfo()
```

# References


